An Open-Source Implementation of the CMPS Algorithm for Assessing Similarity of Bullets

### 2022-04-28 rj_format/rj-format-ju-hofmann/ju-hofmann.Rmd
update rjtools format
added acknowledgement

---
title: "An Open-Source Implementation of the CMPS Algorithm for Assessing Similarity of Bullets"
header-includes:
  - \usepackage{amsmath}
  - \usepackage{xcolor}
  - \usepackage{longtable}
  - \usepackage{adjustbox}            
  - \usepackage[font=small,skip=5pt]{caption}
  - \usepackage{subcaption}
  - \usepackage{afterpage}
  - \usepackage{bbm}
author:
  # see ?rjournal_article for more information
  - name: Wangqian Ju
    affiliation: Department of Statistics
    address:
    - Center for Statistics and Applications in Forensic Evidence
    - Iowa State University
    - 2438 Osborn Dr
    - Ames, IA 50011
    url: https://github.com/willju-wangqian
    email:  wju@iastate.edu
    orcid: 0000-0002-9977-377X
  - name: Heike Hofmann
    url: https://github.com/heike
    email: hofmann@iastate.edu
    orcid: 0000-0002-9079-593X
    affiliation: Department of Statistics
    address:
    - Center for Statistics and Applications in Forensic Evidence
    - Iowa State University
    - 2438 Osborn Dr
    - Ames, IA 50011
abstract: >
  In this paper, we introduce the R package \pkg{cmpsR}, an open-source implementation of the Congruent Matching Profile Segments (CMPS) method developed at the National Institute of Standards and Technology (NIST) for objective comparison of striated tool marks. The functionality of the package is showcased by examples of bullet signatures that come with the package. Graphing tools are implemented in the package as well for users to assess and understand the CMPS results. Initial tests were performed on bullet signatures generated from two sets of 3D scans in the Hamby study under the framework suggested by the R package \pkg{bulletxtrctr}. New metrics based on CMPS scores are introduced and compared with existing metrics. A measure called Sum of squares ratio is included, and how it can be used for evaluating different scans, metrics, or parameters is showcased with the Hamby study data sets. An open-source implementation of the CMPS algorithm makes the algorithm more accessible, generates reproducible results, and facilitates further studies of the algorithm such as methods comparison.
thanks: |
  This work was partially funded by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreement 70NANB20H019 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, Duke University, University of California Irvine, University of Virginia, West Virginia University, University of Pennsylvania, Swarthmore College and University of Nebraska, Lincoln.    
preamble: |
  % Any extra LaTeX you need in the preamble
bibliography: ju-hofmann.bib
type: package
output: 
  rjtools::rjournal_web_article
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: sentence
---

@@

@@

-   `hamby*_result_generator.R`: these R scripts take the processed data of Hamby set 252 and Hamby set 44, generate preliminary results used in the paper, and save the results in the `.csv` format in the folder `data-csv`. Please make sure that the package versions of `bulletxtrctr` and `cmpsR` meet the requirements.
-   `rds_generator.R`: this R script takes the generated `.csv` files and produce `CMPSpaper_results.rds`. `CMPSpaper_results.rds` is identical to `data/CMPSpaper_results.rds` and is used generate figures and other results presented in the paper.

>>>>

---
title: "An Open-Source Implementation of the CMPS Algorithm for Assessing Similarity of Bullets"
header-includes:
  - \usepackage{amsmath}
  - \usepackage{xcolor}
  - \usepackage{longtable}
  - \usepackage{adjustbox}            
  - \usepackage[font=small,skip=5pt]{caption}
  - \usepackage{subcaption}
  - \usepackage{afterpage}
  - \usepackage{bbm}
author:
  # see ?rjournal_article for more information
  - name: Wangqian Ju
    affiliation: Department of Statistics
    address:
    - Center for Statistics and Applications in Forensic Evidence
    - Iowa State University
    - 2438 Osborn Dr
    - Ames, IA 50011
    url: https://github.com/willju-wangqian
    email:  wju@iastate.edu
    orcid: 0000-0002-9977-377X
  - name: Heike Hofmann
    url: https://github.com/heike
    email: hofmann@iastate.edu
    orcid: 0000-0002-9079-593X
    affiliation: Department of Statistics
    address:
    - Center for Statistics and Applications in Forensic Evidence
    - Iowa State University
    - 2438 Osborn Dr
    - Ames, IA 50011
abstract: >
  In this paper, we introduce the R package \pkg{cmpsR}, an open-source implementation of the Congruent Matching Profile Segments (CMPS) method developed at the National Institute of Standards and Technology (NIST) for objective comparison of striated tool marks. The functionality of the package is showcased by examples of bullet signatures that come with the package. Graphing tools are implemented in the package as well for users to assess and understand the CMPS results. Initial tests were performed on bullet signatures generated from two sets of 3D scans in the Hamby study under the framework suggested by the R package \pkg{bulletxtrctr}. New metrics based on CMPS scores are introduced and compared with existing metrics. A measure called Sum of squares ratio is included, and how it can be used for evaluating different scans, metrics, or parameters is showcased with the Hamby study data sets. An open-source implementation of the CMPS algorithm makes the algorithm more accessible, generates reproducible results, and facilitates further studies of the algorithm such as methods comparison.
preamble: |
  % Any extra LaTeX you need in the preamble
bibliography: ju-hofmann.bib
type: package
output: 
  rjtools::rjournal_web_article
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: sentence
---

@@

## Acknowledgement

This work was funded (or partially funded) by the Center for Statistics and Applications in Forensic Evidence (CSAFE) through Cooperative Agreements 70NANB15H176 and 70NANB20H019 between NIST and Iowa State University, which includes activities carried out at Carnegie Mellon University, Duke University, University of California Irvine, University of Virginia, West Virginia University, University of Pennsylvania, Swarthmore College and University of Nebraska, Lincoln.

@@

-   `hamby*_result_generator.R`: these R scripts take the processed data of Hamby set 252 and Hamby set 44, generate preliminary results used in the paper, and save the results in the `.csv` format in the folder `data-csv`. Please make sure that the package versions of \pkg{bulletxtrctr} and `cmpsR` meet the requirements.
-   `rds_generator.R`: this R script takes the generated `.csv` files and produce `CMPSpaper_results.rds`. `CMPSpaper_results.rds` is identical to `data/CMPSpaper_results.rds` and is used generate figures and other results presented in the paper.

### 2022-04-28 rj_format/rj-format-ju-hofmann/ju-hofmann.tex
update rjtools format
added acknowledgement

@@

\begin{itemize}
\tightlist
\item
  \texttt{hamby*\_result\_generator.R}: these R scripts take the
  processed data of Hamby set 252 and Hamby set 44, generate preliminary
  results used in the paper, and save the results in the \texttt{.csv}
  format in the folder \texttt{data-csv}. Please make sure that the
  package versions of \texttt{bulletxtrctr} and \texttt{cmpsR} meet the
  requirements.
\item
  \texttt{rds\_generator.R}: this R script takes the generated
  \texttt{.csv} files and produce \texttt{CMPSpaper\_results.rds}.
  \texttt{CMPSpaper\_results.rds} is identical to
  \texttt{data/CMPSpaper\_results.rds} and is used generate figures and
  other results presented in the paper.
\end{itemize}

>>>>

\hypertarget{acknowledgement}{%
\subsection{Acknowledgement}\label{acknowledgement}}

This work was funded (or partially funded) by the Center for Statistics
and Applications in Forensic Evidence (CSAFE) through Cooperative
Agreements 70NANB15H176 and 70NANB20H019 between NIST and Iowa State
University, which includes activities carried out at Carnegie Mellon
University, Duke University, University of California Irvine, University
of Virginia, West Virginia University, University of Pennsylvania,
Swarthmore College and University of Nebraska, Lincoln.

@@

\begin{itemize}
\tightlist
\item
  \texttt{hamby*\_result\_generator.R}: these R scripts take the
  processed data of Hamby set 252 and Hamby set 44, generate preliminary
  results used in the paper, and save the results in the \texttt{.csv}
  format in the folder \texttt{data-csv}. Please make sure that the
  package versions of \pkg{bulletxtrctr} and \texttt{cmpsR} meet the
  requirements.
\item
  \texttt{rds\_generator.R}: this R script takes the generated
  \texttt{.csv} files and produce \texttt{CMPSpaper\_results.rds}.
  \texttt{CMPSpaper\_results.rds} is identical to
  \texttt{data/CMPSpaper\_results.rds} and is used generate figures and
  other results presented in the paper.
\end{itemize}
